---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# galisats

<!-- badges: start -->

[![R-CMD-check](https://github.com/LechJaszowski/galilean_satellites/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/LechJaszowski/galilean_satellites/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

If you are looking at Jupiter through binoculars or a telescope and don't know which moon is which, then use this package.

`galisats` is used to determine the positions of the four greatest satellites of Jupiter (called Galilean satellites). Positions are shown on the plot for any given time (ET – Ephemeris Time) with respect to the planet, as seen from the Earth.

The `galsat()` function calculates numerical values of the satellites’ positions:

x – the apparent rectangular coordinate of the satellite with respect to the center of Jupiter’s disk in the equatorial plane in the units of Jupiter’s equatorial radius; X is positive toward the west

y – the apparent rectangular coordinate of the satellite with respect to the center of Jupiter’s disk from the equatorial plane in the units of Jupiter’s equatorial radius; Y is positive toward the north

The function is based on algorithms in the book:

Astronomical Formulae for Calculators (4th edition), Jean Meeus, Willmann-Bell Inc., 1988

## Installation

You can install the development version of galisats from [GitHub] (<https://github.com/>) with:

``` r
# install.packages("devtools")
devtools::install_github("LechJaszowski/galilean_satellites")
```

## Example

This is an example of using the function galsat():

```{r example}
library(galisats)
galsat(2025, 10, 13, 21, 40)
```
