---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# galisats

<!-- badges: start -->

<!-- badges: end -->

galisats is used to determine the positions of the four greatest satellites of Jupiter (called Galilean satellites). Positions are calculated for any given time (ET – Ephemeris Time) with respect to the planet, as seen from the Earth.

The galsat() function displays numerical values of the satellites’ positions:

x – the apparent rectangular coordinate of the satellite with respect to the center of Jupiter’s disk in the equatorial plane in the units of Jupiter’s equatorial radius; X is positive toward the west

y – the apparent rectangular coordinate of the satellite with respect to the center of Jupiter’s disk from the equatorial plane in the units of Jupiter’s equatorial radius; Y is positive toward the north

The function is based on algorithms in the book:

Astronomical Formulae for Calculators (4th edition), Jean Meeus, Willmann-Bell Inc., 1988

## Installation

You can install the development version of galisats from [GitHub] (<https://github.com/>) with:

``` r
# install.packages("galisats")
# devtools::install_github("...")
```

## Example

This is an example of using the function galsat():

```{r example}
library(galisats)
galsat(2025, 10, 24, 21, 40)
```
